---
title: "---"
author: "Hashdevrunner"
output: html_document
fontsize: 10
---
```{r echo=FALSE}
library(dplyr); library(reshape2)
```
##Executive Summary
The ```ToothGrowth``` dataset describes the effect of Ascorbic acid in the form of Vitamin C and Orange Juice on Tooth Growth in Guinea Pigs. Experiment showed that there is a significant increase in length on Guinea Pig's toothgrowth as dosage is increased.

##Exploring data
```{r}
data(ToothGrowth)
tg <- ToothGrowth
levels(tg$supp) <- c("Orange Juice","Vitamin C")
tg$dose <- as.factor(tg$dose)
```

Highlighting the basic features of the data, it has 3 sets of subjects for each dosage on the 2 supplements.
```{r}
summary(tg)
table(tg$supp, tg$dose)
```

The boxplot indicates that there is an increasing trend when administering higher dosage of Orange Juice or Vitamin C.
```{r}
library(ggplot2)
plot <- ggplot(tg, 
               aes(x=factor(dose),y=len,fill=factor(dose)))
plot <-plot + geom_boxplot(notch=F) + 
      facet_grid(.~supp) +
       scale_x_discrete("Dosage (mg)") + 
       scale_y_continuous("Length") +  
       ggtitle("Exploratory Data Analyses") + 
       theme(legend.position = "none")

print(plot)
```

##Statistical Inference and Analysis
In a first step, it is natural to verify whether the dosage of the supplement affects the toothgrowth irregardless of the supplement. Since we have 3 dose sets, we will compare the effects of .5mg to 1.0mg dose, 1.0mg to 2.0 dose, and .5mg to 2.0 dose.

```{r}

t1 <- filter(tg, dose == "0.5")$len
t2 <- filter(tg, dose == "1")$len
t3 <- filter(tg, dose == "2")$len

t1vt2 <- t.test(t1,t2, paired = FALSE, var.equal = FALSE, alternative="two.sided")
```
The test to compare .5 and 1.0mg dose rejects the null hypothesis. It leads us to accept the alternative hyposthesis that their means is not equal to 0 and infer that these two samples are different.